### 2023-03-27 temp_code/MS_isoviz_code.Rmd
MS junc to iso code

---
title: "Megan's code for Isoviz"
output: html_notebook
---

# 230327
Write a function that maps leafcutter cluster data back to isoforms 
- use psl/gtf, advantage of psl is it is compatible with long read generated gtf which lacks UTR info, it will likely become increasingly common to use long read data to determine which isoforms are present in a specific cell type
- input gene name and cell type, start with RBFOX2 and A375
- output is clusters with counts and linked isoforms
- print image filtered for cluster with <=4 junctions but print pdf with all
- if junctions doesn't map back, have a parameter to include/not include as novel
- go back and only plot isoforms with junction supporting data
- eventually also want to call Andrews predictions and get best predicted guide for junction
- could we even have a screen design setting? to scale it up?

```{r}
# libraries
require(readr)
require(dplyr)
require(tidyr)
require(magrittr)
require(ggplot2)
require(data.table)
require(stringr)
require(ggpubr)
require(grid)
require(gridExtra)
require(cowplot)

```

```{r}
# generate leafcutter cluster files and store for standard cell types
# this is publicly available A375 data used to design the screen

# get leafcutter data in proper format
clusters <- read_tsv("~/isoviz/data-raw/SRR6515912_out-regtools_.junc.a375_m30_p0.01_.sorted", col_names = FALSE, skip = 1) # 15,901
clusters %<>% 
  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
  select(-same, -extra)

# remove clusters with no counts, not sure why these are here to begin with- only 7
# filter out chrM and chrY clusters
clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM") # 15,886

leafcutter_clusters = as.data.table(clusters)

save(leafcutter_clusters, file = "~/isoviz/data/leafcutter_clusters.rda")
```

```{r}
# could make this a function
# create gene name conversion file from gtf
gtf = read_tsv("~/isoviz/data-raw/gencode.v41.basic.annotation.gtf", col_names = FALSE, skip = 5) %>% filter(X3 == "transcript")
gtf$transcript_id = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[2]])
gtf$transcript_id = sapply(gtf$transcript_id, gsub, pattern = "transcript_id", replacement = "")
gtf$transcript_id = sapply(as.character(gtf$transcript_id), gsub, pattern = '"', replacement = '')
gtf$transcript_id = sapply(as.character(gtf$transcript_id), gsub, pattern = ' ', replacement = '')

gtf$gene_id = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[1]])
gtf$gene_id = sapply(gtf$gene_id, gsub, pattern = "gene_id", replacement = "")
gtf$gene_id = sapply(as.character(gtf$gene_id), gsub, pattern = '"', replacement = '')
gtf$gene_id = sapply(as.character(gtf$gene_id), gsub, pattern = ' ', replacement = '')

gtf$transcript_name = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[6]])
gtf$transcript_name = sapply(gtf$transcript_name, gsub, pattern = "transcript_name", replacement = "")
gtf$transcript_name = sapply(as.character(gtf$transcript_name), gsub, pattern = '"', replacement = '')
gtf$transcript_name = sapply(as.character(gtf$transcript_name), gsub, pattern = ' ', replacement = '')

gtf$gene_name = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[4]])
gtf$gene_name = sapply(gtf$gene_name, gsub, pattern = "gene_name", replacement = "")
gtf$gene_name = sapply(as.character(gtf$gene_name), gsub, pattern = '"', replacement = '')
gtf$gene_name = sapply(as.character(gtf$gene_name), gsub, pattern = ' ', replacement = '')

gtf$gene_type = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[3]])
gtf$gene_type = sapply(gtf$gene_type, gsub, pattern = "gene_type", replacement = "")
gtf$gene_type = sapply(as.character(gtf$gene_type), gsub, pattern = '"', replacement = '')
gtf$gene_type = sapply(as.character(gtf$gene_type), gsub, pattern = ' ', replacement = '')

gtf$transcript_type = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[5]])
gtf$transcript_type = sapply(gtf$transcript_type, gsub, pattern = "transcript_type", replacement = "")
gtf$transcript_type = sapply(as.character(gtf$transcript_type), gsub, pattern = '"', replacement = '')
gtf$transcript_type = sapply(as.character(gtf$transcript_type), gsub, pattern = ' ', replacement = '')

gtf %>% dplyr::select(gene_id, trans_id = transcript_id, gene_name, transcript_name, gene_type, transcript_type) %>% distinct() %>%
  write_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)

```

```{r}
# modify psl file data to include other gene and transcript names
# currently using gencode basic anntotation file
convert = read_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)

load("~/isoviz/data/iso_exon_data.rda") # 839,796
iso_data_complete = iso_exon_data %>% left_join(convert, by = c("gene_id", "trans_id"))

save(iso_data_complete, file = "~/isoviz/data/iso_data_complete.rda")

```

```{r}
# make psl with intron coords instead of exon
# adapted from genomedata.R

mydataset <- fread("~/isoviz/data-raw/gencode.v41.basic.annotation.psl")

mydataset$gene_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][2]})
mydataset$trans_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][1]})

# Get coordinates for each "block"
mydataset$blocksizes = sapply(mydataset$V19, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
mydataset$blockstarts = sapply(mydataset$V21, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
mydataset$strand = mydataset$V9
mydataset$chr = mydataset$V14
mydataset$start = mydataset$V16
mydataset$end = mydataset$V17
mydataset$transcript_length = mydataset$end - mydataset$start

# Clean up and seperate blocks into rows
# need to do different adjustments here to get intron coords that will line up with leafcutter
mydataset = mydataset %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
                            transcript_length) %>% separate_rows(blocksizes, blockstarts)
mydataset$blockstarts=as.numeric(mydataset$blockstarts)
mydataset$blocksizes=as.numeric(mydataset$blocksizes)
mydataset$blockends = mydataset$blockstarts + mydataset$blocksizes

transcript_ids = unique(mydataset$trans_id)
length(transcript_ids) # 116,566
trans_id = c()
intron_starts = c()
intron_ends = c()

for(id in transcript_ids){
  data = mydataset %>% filter(trans_id == id)
  new_ends = data$blockstarts[-1] + 1
  len = nrow(data)
  new_starts = data$blockends[-len]
  n = len-1
  trans_id = c(trans_id, rep(id, n))
  intron_starts = c(intron_starts, new_starts)
  intron_ends = c(intron_ends, new_ends)
}

# Important note: this no longer has single exon transcripts
intron_data = data.frame(trans_id, intron_starts, intron_ends) # 723,230

convert = read_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)
trans_info = mydataset %>% select(1, 4, 5, 6) %>% distinct() %>% filter(chr != "chrY", chr != "chrM")

data = intron_data %>% left_join(trans_info,  by = "trans_id") %>% left_join(convert, by = c("gene_id", "trans_id")) %>% select(4, 2, 3, 5, 1, 6, 7, 8, 9, 10)

iso_intron_data = as.data.table(data)

save(iso_intron_data, file = "~/isoviz/data/iso_intron_data.rda")

```

```{r}
load(file = "~/isoviz/data/iso_intron_data.rda")
load(file = "~/isoviz/data/leafcutter_clusters.rda")

# map junctions function- for now, it isn't using cell type but eventually we should change this
map_junction = function(gene = "ENSG00000001167.15", cell_type = "A375"){

  # filter for gene, for now, also filter for protein coding transcripts- can set this as a parameter later
  gene_iso_data = iso_intron_data %>% filter(grepl(gene, gene_id), transcript_type == "protein_coding")

  # join iso data with leafcutter data and mark any unannotated junctions
  gene_cluster = gene_iso_data %>% left_join(leafcutter_clusters, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>% filter(!is.na(cluster.n))
  gene = unique(gene_cluster$gene_id)
  name = unique(gene_cluster$gene_name)
  gene_strand = unique(gene_cluster$strand)
  cluster_ids = unique(gene_cluster$cluster.n)
  output = leafcutter_clusters %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
    left_join(gene_iso_data, by = c("chr", "start" = "intron_starts", "end" = "intron_ends")) %>% 
    mutate(junc.usage = round((junc.counts/cluster.counts)*100, digits = 0), 
           transcript_name = ifelse(is.na(transcript_name), "Unannotated", transcript_name),
           trans_id = ifelse(is.na(trans_id), "Unannotated", trans_id),
           gene_id = ifelse(is.na(gene_id), gene, gene_id),
           gene_name = ifelse(is.na(gene_name), name, gene_name),
           strand = ifelse(is.na(strand), gene_strand, strand))

  return(output)
}

to_plot = map_junction("ENSG00000001167.15") # NFYA
to_plot = map_junction("ENSG00000100320") # RBFOX2

```

```{r}
load(file='~/isoviz/data/iso_data_complete.rda')
to_plot

# plot isoform plus leafcutter clusters
plot_junction_to_isoform = function(to_plot){

  # get info
  gene = unique(to_plot$gene_id)
  name = unique(to_plot$gene_name)
  transcripts = unique(to_plot$transcript_name)

  # I am copying the plot_isoforms function here, figure out how to integrate these later
  gene_trans =which(str_detect(iso_data_complete$gene_id, gene))
  gene_data = iso_data_complete[gene_trans,]

  # filter for transcripts that have junction level data to support
  gene_data %<>% filter(transcript_name %in% transcripts)

  # setting parameters for the plot based on length of the gene and number of isoforms
  length = max(gene_data$end) - min(gene_data$start)
  min_start = min(gene_data$start)
  n = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
  y_max = n + 1
  max_end = max(gene_data$end)

  # order the isoforms by length for plotting
  order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
    dplyr::arrange(desc(transcript_length), desc(n))
  order_plot$trans_order = 1:nrow(order_plot)
  order_plot = order_plot %>% dplyr::select(transcript_name, trans_order)

  to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
    arrange(trans_order, blockstarts) %>%
    mutate(y1 = trans_order - 0.2, y2 = trans_order + 0.2)
  to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)

  # make isoform plot
  p1 = ggplot() +
    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = y1, ymax = y2)) +
    xlim(min_start, max_end) + ylim(0,y_max) +
    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name)) +
    theme_bw() + ylab("") + xlab("") +
    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top") +
    ggtitle(paste0(name, " (", to_plot_ordered$strand[1], ")")) +
    theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5)) + coord_cartesian(xlim = c(min_start, max_end))
    #coord_cartesian(clip = 'off')
    #theme(plot.margin = margin(0.5,1.5,0.5,0.5, "in"))

  # new code to plot leafcutter cluster info underneath
  clusters = unique(to_plot$cluster.n)
  introns = to_plot %>% select(-gene_id, -trans_id, -gene_name, -transcript_name) %>% distinct() %>% group_by(cluster.n) %>% mutate(junc_order = row_number())
  p2 = ggplot() + geom_rect(data = introns, mapping = aes(xmin = start, xmax = end, ymin = junc_order - 0.2, ymax = junc_order + 0.2)) + 
    xlim(min_start, max_end) + theme_bw() + facet_wrap(~ cluster.n, scales = "free_y", ncol = 1, strip.position = "top") + 
    xlab("Hg38 Genomic Position (bp)") + ylab("") + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
    coord_cartesian(xlim = c(min_start, max_end))

  plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(3, 3))
  ggsave("~/isoviz/plots/rbfox2_isoform_draft.pdf", width = 8, height = 4)

    # try using pivot for aggregating isoform

}
```